---
title: "01 - ggplot2 intro"
author: "Ashwin Malshe"
date: "6/3/2021"
output: 
  html_document:
    theme: "paper"
---

The available themes for rmarkdown are as follows:

default, cerulean, journal, flatly, darkly, readable, spacelab, united, cosmo, lumen, paper, sandstone, simplex, yeti

You can check out how they look by visiting this website: https://bootswatch.com

Also, if you want to use a theme that is not available in rmarkdown, simply download the CSS file and save it in the same folder where you saved the rmarkdown file. Next, replace `theme: paper` in the YAML header above with `css: FILE_NAME.css` where `FILE_NAME.css` is the CSS file you downloaded.


```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE, cache.extra = packageVersion('tint'))

library(ggplot2)
library(dplyr)
```


## Plot `mpg` data set

`mpg` or miles per gallon data set comes with the R installation.

```{r}
head(mpg)
```

```{r}
summary(mpg)
```

```{r}
str(mpg)
```



Let's begin by creating an empty plot.


```{r}
ggplot()
```


## Aesthetics

Once we have an empty plot, we can start adding layers to it. `ggplot2` requires a data frame. It can't work with matrix or list or other data types.


Our X axis is `displ` and Y axis is `cty`. These are aethetics *mappings* for `ggplot2`. They are wrapped inside an `aes()` function.


```{r}
ggplot(data = mpg,
       mapping = aes(x = displ, y = cty))
```

Now, we mapped X and Y axes to two variables in the data but no plot is visible. This is because we need to provide this instruction to `ggplot2` using `geom_*()` functions. These are geometric shapes for `ggplot2`.

As both `displ` and `cty` are numeric variables, it makes sense to create a scatterplot.

```{r}
ggplot(data = mpg,
       mapping = aes(x = displ, y = cty)) +
  geom_point() # this layer adds points to the plot
```


And just like that you created your first scatterplot!

As a data scientist, I am tempted to overlay a line on this scatterplot to summarize the relationship. We can acheive that by using `geom_smooth()`

```{r}
ggplot(data = mpg,
       mapping = aes(x = displ, y = cty)) +
  geom_point() +
  geom_smooth()
```

By default, ggplot2 adds a line estimated using LOESS (Savitzky–Golay filter) (https://en.wikipedia.org/wiki/Local_regression)

We can easily change it to a linear regression by specifying `method` argument.

```{r message = FALSE}
ggplot(data = mpg,
       mapping = aes(x = displ, y = cty)) +
  geom_point() +
  geom_smooth(method = "lm")
```


## Boxplot



```{r}
ggplot(data = mpg,
       mapping = aes(y = cty)) +
  geom_boxplot()

```



```{r}
ggplot(data = mpg,
       mapping = aes(x = manufacturer, y = cty)) +
  geom_boxplot()

```

## Bar plot

```{r}
table(mpg$class)
```

```{r}
ggplot(data = mpg,
       mapping = aes(x = class)) +
  geom_bar()
```


```{r}
dplyr::count(mpg, class)
```


```{r}
ggplot(data = mpg,
       mapping = aes(x = class)) +
  geom_bar(fill = "blue", # sets the color inside the bars
           alpha = 0.5, # sets the transparency
           color = "red", # sets the border color. The default is black
           size = 1 #sets the line thickness
           ) 
```
### Aside

`geom_bar()` internally creates a frequency distribution table and uses it to create a bar plot.

However, often we may already have a data set summarized as a frequency distribution. In that case, use `stat = identity` argument inside `geom_bar()`

```{r}
class_freq <- dplyr::count(mpg, class)
```

```{r}
ggplot(data = class_freq,
       mapping = aes(x = class, y = n)) +
  geom_bar(stat = "identity")
```


Another alternative is to use a wrapper for `geom_bar(stat = "identity")` named `geom_col()`

```{r}
ggplot(data = class_freq,
       mapping = aes(x = class, y = n)) +
  geom_col() # This is equal to geom_bar(stat = "identity")
```


## Histogram

Bar graph for a continuous variable


```{r}
ggplot(data = mpg,
       mapping = aes(x = cty)) +
  geom_histogram(bins = 12,
                 fill = "navy",
                 color = "white")
```

## Jitter plot

When the data has many similar observations, scatterplot can be misleading because it may have many overlapping points. One way to handle that is to make points more transparent.

```{r}
ggplot(data = mpg,
       mapping = aes(x = manufacturer, y = cty)) +
  geom_point(alpha = 0.2)

```
Another option is to add some random noise to the points to create a jittering effect

```{r}
set.seed(999)
ggplot(data = mpg,
       mapping = aes(x = class, 
                     y = cty,
                     color = class)) +
  geom_jitter()

```

You can also add a violin plot to this. Violin plot shows you the distribution of the data subset.

```{r}
set.seed(999)
ggplot(data = mpg,
       mapping = aes(x = class, 
                     y = cty,
                     color = class)) +
  geom_violin() +
  geom_jitter()

```




