---
title: "Ready made themes"
subtitle: "Demo using ggthemes and hrbrthemes"
author: "Ashwin Malshe"
date: "`r Sys.Date()`"
output: tint::tintHtml

---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))

library(tidyverse)
library(ggthemes)
library(hrbrthemes)

```

**Note that I am using a different theme for rmarkdown. This theme is based on Tufte style.**

## Give me some style! Using readymade themes

`ggplot2` allows fine-grain tuning of the looks of the plot. However, often we may be content with ready made themes. R ecosystem has created numerous themes and made them available through packages. I strongly recommend `ggthemes` and `hrbrthemes` packages. [`hrbrthemes` is written by Bob Rudis, who is a cyber security expert.]

Let's create a simple base plot.

```{r}
base <- ggplot(mpg, aes(cty, hwy)) +
  geom_point(aes(fill = manufacturer), 
             color = "white", 
             shape = 21, # this shape allows you to choose border color for the point
             size = 3)

base
```

Now we will modify this base plot using different themes. Let's start with `hrbrthemes`.

## `hrbrthemes`



```{r}
base +
  theme_ipsum_inter()
```

```{r}
base +
  theme_ipsum_rc()
```
```{r}
base +
  theme_ipsum_es()
```



```{r}
base +
  theme_modern_rc()
```


## `ggthemes`

`ggthemes` packs a lot of themes based on famous websites and publishers such as The Economist, and fivethirtyeight blog.

```{r}
base +
  theme_economist()
```

```{r}
base +
  theme_fivethirtyeight()
```

```{r}
base +
  theme_hc()
```

```{r}
base +
  theme_solarized()
```
