---
title: "Scales, Axes, and Legends"
author: "Ashwin Malshe"
date: "06/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Basics of scales

```{r}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(color = class))

```

This is what ggplot2 is actually plotting. You can change the scales if you want. For example, you can drop the X axis label by writing NULL inside the parenthesis linked to scale_x_continuous

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete()
```

Although you can have many `geom_*()` layers of the same `geom` in your `ggplot`, you can only have one `scale_*()` function of that specific type. For example, is you used `scale_x_continuous()` once, you should avoid using it again. If you use it more than once, `ggplot2` will use the **last** `scale_x_continuous()`. It will ignore all its occurences before that.

Simplify the following code:

```{r}
ggplot(mpg, aes(y = displ, x = class)) +
  scale_y_continuous("Displacement (l)") +
  scale_x_discrete("Car type") +
  scale_x_discrete("Type of car") +
  scale_colour_discrete() +
  geom_point(aes(colour = drv)) +
  scale_colour_discrete("Drive\ntrain")
```

# Guides

Now we will learn how to alter the guides - legends and axes


## Change axis and legend titles

```{r}
ggplot(mpg,aes(displ, hwy, color = class)) +
  geom_point()
```

```{r}
ggplot(mpg,aes(displ,hwy, color = class)) +
  geom_point() +
  scale_x_continuous("Displacement") +
  scale_y_continuous("Highway Miles")
```


Change the legend title

```{r}
ggplot(mpg,aes(displ,hwy, color = class)) +
  geom_point() +
  scale_x_continuous("Displacement") +
  scale_y_continuous("Highway Miles") +
  scale_color_discrete("Vehicle Class")
```

Multiline legend title
```{r}
ggplot(mpg,aes(displ,hwy, color = class)) +
  geom_point() +
  scale_x_continuous("Displacement") +
  scale_y_continuous("Highway Miles") +
  scale_color_discrete("Vehicle\nClass")
```

Use a wrapper 

```{r}
ggplot(mpg,aes(displ,hwy, color = class)) +
  geom_point() +
  labs(x = "Displacement", y = "Highway Miles", color = "Vehicle\nClass")
```

## Ticks and breaks

```{r}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point() +
   scale_x_continuous(breaks = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5))
```


```{r}
ggplot(mpg, aes(displ, hwy, color = class)) +
  geom_point() +
  scale_x_continuous(breaks = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5),
                     labels = c("1.5 lit.", "B", "C", "D", "E", "F"))
```

Change the legend key labels

```{r}
ggplot(mpg,aes(displ,hwy, color = class)) +
  geom_point() +
  scale_color_discrete(labels = c("1", "2", "3", "4", "5", "6", "7")) +
  scale_x_continuous(limits = c(0, 7),
                     breaks = c(0:7))
```

Remove the titles and labels from legends and additionally breaks from axes

```{r}
ggplot(mpg,aes(displ,hwy, color = class)) +
  geom_point() +
  scale_x_continuous(NULL,breaks = NULL,labels = NULL) +
  scale_y_continuous(NULL,breaks = NULL,labels = NULL) +
  scale_color_discrete(NULL, labels = NULL)
```

## Using scales package to change number formatting on axes and legends

```{r}

set.seed(123456)
d1 <- diamonds[sample(nrow(diamonds), 100),]

ggplot(d1, aes(clarity, price)) +
  geom_point(alpha = 1/2)
```

The y axis has prices, which may need two tweaks. First, there is no comma separating
thousands. Second, we may want to show units ($).

```{r}
ggplot(d1, aes(clarity,price)) +
  geom_point(alpha = 1/2) +
  scale_y_continuous(labels = scales::comma_format())
```

Add $ sign before the price. You can add any currency symbol here.

```{r}
ggplot(d1, aes(clarity, price)) +
  geom_point(alpha = 1/2) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$"))
```

**We won't do the code below this in today's session. Please check it out at your convenience.**

## Controlling legend position using theme(). We will learn more about themes in the next session.

```{r}
ggplot(mydata, aes(p, q)) +
  geom_point(aes(color = r), size = 4) +
  ggtitle("Legends at the top") +
  theme(legend.position = "top") 
```

```{r}
ggplot(mydata, aes(p, q)) +
  geom_point(aes(color = r), size = 4) +
  ggtitle("Legends at the bottom") +
  theme(legend.position = "bottom")
```

```{r}
ggplot(mydata, aes(p,q)) +
  geom_point(aes(color = r), size = 4) +
  ggtitle("Legends on left") +
  theme(legend.position = "left")
```

```{r}
ggplot(mydata, aes(p,q)) +
  geom_point(aes(color = r), size = 4) +
  ggtitle("No legends") +
  theme(legend.position = "none")
```


You can bring the legends inside the plot as well.

```{r}
ggplot(mydata, aes(p,q)) +
  geom_point(aes(color = r), size = 4) +
  theme(legend.position = c(1,0.5))
```

```{r}
ggplot(mydata, aes(p,q)) +
  geom_point(aes(color = r), size = 4) +
  theme(legend.position = c(0.9, 0.5), 
        legend.margin = margin(t = 0,
                               r = 0,
                               b = 0,
                               l = 0,
                               "mm")) +
  scale_color_discrete(NULL, labels = NULL)
```

Can you remove the leftover white patch?

You can also change the direction in which legends are laid out

```{r}
ggplot(mydata, aes(p, q)) +
  geom_point(aes(color = r), size = 4) +
  theme(legend.direction = "horizontal")

```

## Guides
 Guides are the legend keys. They are shown in a table and therefore you can alter the table rows and columns.
 
 
```{r}
ggplot(d1, aes(carat, price)) +
  geom_point(aes(color = clarity)) +
  guides(color = guide_legend(ncol = 3))
```

The default for filling up the guides is by column. You can also fill them by row.

```{r}
ggplot(d1, aes(carat, price)) +
  geom_point(aes(color = clarity)) +
  guides(color = guide_legend(ncol = 3, byrow = TRUE))
```

Reverse the ordering of legend keys. Note that this reverses simply the order in which the keys are displayed and not the actual scaling. For example, IF has color pink in both the cases. The reversal simply displays IF first instead of last in the legend.

```{r}
ggplot(d1, aes(carat, price)) +
  geom_point(aes(color = clarity)) +
  guides(color = guide_legend(ncol = 3, byrow = TRUE, reverse = TRUE))
```

Finally, what if you want to override the aesthetics in any geom and display legends with its own aes. This is useful when we have cases where we used a small alpha to reduce overplotting but we want the legend to have darker colors.

```{r}
ggplot(d1, aes(carat, price)) +
  geom_point(aes(color = clarity), alpha = 0.4)
```

We can override this alpha in guides

```{r}
ggplot(d1, aes(carat, price)) +
  geom_point(aes(color = clarity), alpha = 0.4) +
  guides(color = guide_legend(override.aes = list(alpha = 1)))
```

Similar to guide_legend, another function guide_colorbar controls the behavior of the guide when we are using a contonuous variable. 

```{r}
mydata2 <- data.frame(a = 1:3, b = 1:3, c = 1:3)
ggplot(mydata2, aes(a,b)) +
  geom_point(aes(color = c))

```

You can use similar options in guide_colorbar too

```{r}
ggplot(mydata2, aes(a,b)) +
  geom_point(aes(color = c)) +
  guides(color = guide_colorbar(reverse = T))
```

What happens when you use guide_legend on a continous variable?

```{r}
ggplot(mydata2, aes(a,b)) +
  geom_point(aes(color = c)) +
  guides(color = guide_legend())
```

Now if you apply the same logic and use guide_colorbar for a discrete variable will it work? Why or why not?
