---
title: "Facets"
output: html_document
---

# `facet_wrap` and `facet_grid`

In this tutorial we will learn Faceting in `ggplot2`

```{r}
library(ggplot2)
```

We will use only a subset of the `mpg` data set as used in chapter 7 because it's easy for the demo. We will remove cars with 5 cylinders (only 4 observations) and class 2seater (5 observations). We will also remove cars with rear wheel drive (25 observations)

```{r}
table(mpg$cyl)
table(mpg$drv)
table(mpg$class)
```


```{r}
mpg2 <- dplyr::filter(mpg, 
               cyl != 5 & drv %in% c("4", "f") & class != "2seater")

```


##`facet_wrap`

Let's start with `facet_wrap`. It simply wraps facets along horizontal or vertical direction


```{r}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL)
```


```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class)
```

A few things to note here for the default behavior.
1. `class` labels are organized alphabetically
2. `class` labels go from left to right
3. `class` labels start from the top row

All these are changeable if you want. You can change the ordering of the faceting variable by relevling the factor variable.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~relevel(as.factor(mpg2$class), ref = "suv"))
```

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~factor(mpg2$class,levels = c("suv","subcompact","midsize","pickup","compact","minivan")))
```

You can change the way the facets are wrapped by changing `dir` option.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, nrow = 2, dir = "v")
```

Finally, you can control where the facets start showing using `as.table` option, which by default is `TRUE`

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, as.table = FALSE)
```






## `facet_grid`

`facet_grid` differs from `facet_wrap` in that it can create grids with two variables. Thus, unlike `facet_wrap` it's truly 2-D.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_grid(.~class)
```

As you can see, `facet_grid` doesn't automatically create multiple rows. This is because we have not given it a second dimension.

Before, we move into the 2-D plotting, study the following code and note the difference

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_grid(class~.)
```

### How to get a 2-D graph?

Now, let's get a 2-D graph where we want the facets laid down by `drv` and `class`

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL) +
  facet_grid(drv~class)
```


## Changing scales

Let's look into more advanced functionality. Start by plotting points on the graph.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class)
```

If you are using this panel for heloing wiht your modeling, there is nothing much you need to do here. Perhaps you could superimpose a regression line but that's about it. There are cases where you don't want to make comparisons and just study each panel standalone. In that case you can change the scales so that they become free.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, scales = "free")
```

Note that now comparisons across the panels are not meaningful because the scales have changed. You can also control for the scale only on x or y axis using `scales = "free_x` and `scales = "free_y` options respectively.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_point() +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, scales = "free_y")
```


### Why across panels comparison becomes meaningless with free scales?

In order to answer this question, let's plot regression lines on the scatterplots. First consider no free scales and then free y scale and free 
x scale.

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, scales = "fixed")
```

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, scales = "free_y")
```

```{r echo=TRUE}
ggplot(mpg2, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  xlab(NULL) +
  ylab(NULL) +
  facet_wrap(~class, scales = "free_x")
```


Clearly the latter two graphs are misleading in case you are comparing the slopes of the lines across the panels!

## Using layers in facets

This example shows an awesome application from Chapter 7. You may have come across these types of plots in the past.

In the following code, we are generating a data frame with 3 variables. Note that `rnorm()` function has a vector for means. This will generate a variable with a mixture of 3 normal distributions.

```{r}
set.seed(345908)
df <- data.frame(x = rnorm(120, c(0, 2, 4)),
                 y = rnorm(120, c(1, 2, 1)),
                 z = letters[1:3])
```

Now, let's plot the points

```{r}
ggplot(df, aes(x, y)) +
  geom_point(aes(color = z),size = 3) +
  theme_bw()+
  facet_wrap(~z)
```

Next, we will superimpose the distribution of all the points on top of each panel. For this, we will create a second data frame which has no z variable.

```{r}
df2 <- dplyr::select(df, -z)

ggplot(df, aes(x, y)) +
  geom_point(data = df2, colour = "grey70", size = 3) +
  geom_point(size = 4, show.legend = TRUE) + # This is simply to give points an outline
  geom_point(aes(colour = z), size = 3) +
  theme_bw() +
  facet_wrap(~z)
```

What if you want to change the default colors of the points?

For that let's create a palette that we want to use.

```{r}

mycolors <- c("#e7298a","#ff7f00","#7570b3")

```


```{r}

ggplot(df, aes(x, y)) +
  geom_point(data = df2, colour = "grey70", size = 3) +
  geom_point(size = 4, show.legend = TRUE) + # This is simply to give points an outline
  geom_point(aes(colour = z), size = 3) +
  theme_bw() +
  scale_color_manual(values = mycolors) +
  facet_wrap(~z)
```


## Exercises

1. Diamonds: display the distribution of `price` conditional on `cut` and `carat`. Try facetting by `cut` and grouping by `carat`. Try facetting by `carat` and grouping by `cut`. Which do you prefer?

2. Recreate the following plot. It facets `mpg2` by `class`, overlaying a smooth curve ï¬t to the full dataset.


```{r}
ggplot(diamonds,aes(x = price,group=carat)) +
  geom_histogram(aes(fill = carat)) +
  facet_wrap(~cut)

```


The graph doesn't look good because `carat` is a continuous variable. Let's convert it into a discrete variable. I prefer to use quntiles for this example.

```{r}
diamonds2 <- diamonds
diamonds2$carat2 <- cut_number(diamonds2$carat,n=5)
```


```{r}
ggplot(diamonds2,aes(x = price,group=carat2)) +
  geom_histogram(aes(fill = carat2)) +
  facet_wrap(~cut,scales = "free_y")

```

```{r}
ggplot(diamonds2,aes(x = price,group=cut)) +
  geom_histogram(aes(fill = cut)) +
  facet_wrap(~carat2,scales = "free_y")
```

```{r}

ggplot(mpg2,aes(displ, hwy)) +
  geom_point() +
  geom_smooth(data = dplyr::select(mpg2, -class), se = FALSE) +
  facet_wrap(~class)
```

