---
title: "Class code"
author: "Ashwin Malshe"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

Example using `labs()` function

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  labs(x = "Displacement (liters)",
       y = "Highway Mileage (miles/gallon)",
       color = "Class of Vehicle",
       title = "My title",
       subtitle = "My subtitle",
       caption = "Credit goes to Base R
                  © UTSA") +
  theme_minimal()

```



Example of `geom_smooth()` where we add regression lines for each `class` but remove standard error bands using `se = FALSE` argument.



```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point(aes(color = class)) +
  geom_smooth(aes(color = class), 
              method = "lm",
              se = FALSE)
```

The previous graph looks cluttered. Use `facet_wrap()` to split the graph.

```{r}
ggplot(mpg2, aes(displ, cty)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  facet_wrap(~ class)
```

What if we want to create above facets for levels of another variable `drv`. 

## Option 1

First filter `mpg2` for only one level of `drv` and then make a facet plot.


`drv == "4"`
```{r}

mpg2 %>% 
  filter(drv == "4") %>% 
ggplot(aes(displ, cty)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  facet_wrap(~ class)
```


`drv == "f"`

```{r}

mpg2 %>% 
  filter(drv == "f") %>% 
ggplot(aes(displ, cty)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  facet_wrap(~ class)
```


## Option 2

Use `facet_grid()`

```{r}
ggplot(mpg2, aes(displ, cty)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  facet_grid(drv ~ class)
```

## Adding minimum and maximum to the plot of averages


```{r}
hwymean2 <- mpg %>%
            group_by(class) %>%
              summarize(avghwy = mean(hwy),
                        maxhwy = max(hwy),
                        minhwy = min(hwy)) %>%
            ungroup()
```


```{r}

ggplot(mpg, aes(class, hwy)) +
  geom_point() +
  geom_point(aes(y = avghwy), 
             hwymean2, 
             color = "red", 
             size = 4,
             alpha = 0.5) +
  geom_text(aes(label = round(avghwy, 2), y = avghwy), 
            data = hwymean2, 
            color = "red",
            hjust = 1.3) +
  geom_hline(aes(yintercept = avghwy), 
             data = hwymean1, 
             color = "blue") +
  geom_text(aes(x = "compact", y = 22, label = round(avghwy,2)),
            data = hwymean1, 
            color = "blue", 
            hjust = 1.5) +
  # The code so far was what we did in the class
  # Hereon what we added in the class
  geom_point(aes(y = maxhwy), 
             hwymean2, 
             color = "green", 
             size = 4,
             alpha = 0.7) +
  geom_point(aes(y = minhwy), 
             hwymean2, 
             color = "navy", 
             size = 4,
             alpha = 0.7)

```