---
title: "Scales"
author: "Ashwin Malshe"
date: "22 October 2021"
output: html_document
---

Install package `pacman` by running the following command in the console. **Don't run it in any R chunk below!**

install.packages("pacman")


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, lubridate, wesanderson, viridis)
theme_set(theme_minimal())
```

In this code, I will focus on a few common uses of `scale_*_*` functions.


## Changing the axis labels

We will use `economics` data set which comes bundled with `ggplot2`. Here is the variable description:

A data frame with 574 rows and 6 variables:

date - Month of data collection
pce - personal consumption expenditures, in billions of dollars
pop - total population, in thousands
psavert - personal savings rate
uempmed - median duration of unemployment, in weeks
unemploy - number of unemployed in thousands

```{r}
head(economics)
```



```{r}
p1 <- economics %>% 
  ggplot(aes(x = date, y = pce)) +
  geom_line()

p1
```

### Add dollar prefix to the Y axis labels

```{r}
p1 +
  scale_y_continuous(labels = scales::dollar)
```

### Add dollar prefix to the Y axis labels **and** "bil" suffix

```{r}
p1 +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", suffix = " bil"))
```

### Change the breaks on the X axis to 5 year intervals.


```{r}
p1 +
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", suffix = " bil")) +
  scale_x_date(breaks = as.Date(paste0(seq(1970, 2010, 5), "-01-01")),
               labels = lubridate::year)
```

### Change the default color palette

For colors, it is better to use `mpg` data set because it has many categorical variables.
Here I am filtering out observations with `2seater` and `minivan` because the color palettes used in `wesanderson` package has usually 5 distinct colors. So it becomes easier for me to demonstrate the use of the scale function with 5 distinct categories of `class`.

```{r}
p2 <- mpg %>% 
  filter(!class %in% c("2seater", "minivan")) %>% 
  ggplot(aes(x = cty, y = hwy, color = class)) +
  geom_point(size = 4, alpha = 0.7)

p2
```


We will use `scale_color_manual()` to change the colors manually.

### `wesanderson` package 

This package is based on the colors used in Wes Anderson movies: https://github.com/karthik/wesanderson
The package developer Karthik Ram is one of the most productive and talented R developers.

```{r}
p2 +
  scale_color_manual(values = wes_palette("Darjeeling1"))
```

Play with other color palettes.

### viridis color palette

**`viridis` is a collection of colorblind-friendly color palettes**

Its developer [Bob Rudis](https://rud.is) is a cybersecurity expert, but he is also a fantastic R developer with several packages under his belt. Check out the `viridis` vignette: https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html


```{r}
p2 +
  scale_color_manual(values = viridis(5))
```

Bob has also added scale functions to directly use these colors.

```{r}
p2 +
  scale_color_viridis(discrete=TRUE)
```

Pick different palettes using `option` argument inside `scale_color_viridis()`

*option* - A character string indicating the colormap option to use. Four options are available: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").

```{r}
p2 +
  scale_color_viridis(option = "A", discrete=TRUE)
```

Let's take one last example with a continuous color scale.

```{r}
p3 <- mpg %>% 
  count(class) %>% 
  ggplot(aes(x = reorder(class, -n), y = n, fill = n)) +
  geom_col()

p3
  
```
```{r}
p3 +
  scale_fill_viridis(direction = -1)
  
```
