---
title: "Annotations"
author: "Ashwin Malshe"
date: "22 October 2021"
output: 
  html_document:
    fig_retina: 1
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, showtext, jpeg, png)
theme_set(theme_minimal())
```

We will use `economics` data set which comes bundled with `ggplot2`. Here is the variable description:

A data frame with 574 rows and 6 variables:

date - Month of data collection
pce - personal consumption expenditures, in billions of dollars
pop - total population, in thousands
psavert - personal savings rate
uempmed - median duration of unemployment, in weeks
unemploy - number of unemployed in thousands


## Text annotations


```{r}
p1 <- economics %>% 
  mutate(year = lubridate::year(date)) %>% 
  filter(year >= 2010) %>% 
  group_by(year) %>% 
  summarize(psavert_mean = mean(psavert, na.rm = TRUE),
            .groups = "drop") %>% 
  ggplot(aes(x = as.character(year), y = psavert_mean)) +
  geom_col()

p1
```


```{r}
p1 +
  geom_text(aes(label = paste0(round(psavert_mean, 2), "%")))
```


```{r}
p1 +
  geom_text(aes(label = paste0(round(psavert_mean, 2), "%")),
            vjust = -0.3, size = 3)
```
```{r}
p1 +
  geom_text(aes(label = paste0(round(psavert_mean, 2), "%"),
                y = 5), 
            size = 3, color = "white")
```

```{r}
p2 <- economics %>% 
  mutate(unemploy_per = unemploy / pop) %>% 
  ggplot(aes(date, unemploy_per)) +
  geom_line() +
  scale_y_continuous(labels = scales::percent)

p2
```

Annotate 2010 spike in the unemployment

```{r}
p2 +
  annotate("text", label = "Mortgage crisis", 
           x = as.Date("2010-01-01"), y = 0.051,
           color = "red")
```

## Image annotation

For annotating plots with images like `geom_text` we have to update the data set with the file path to those images. It's a bit tedious and you shoudl use it only if you really need to. The package for this is `ggimage` and the relevant geom is `geom_image()`. Read more about it here:
https://yulab-smu.top/pkgdocs/ggimage.html


It's far more common to use images to annotate only one point of a plot. For example, you may want to include a logo on the plot either in the background or in one corner. Here is how to do that:

Download UTSA logo (stacked) from here and save it in your project directory.

https://www.utsa.edu/marcomstudio/resources/logos-visual-identity/

Take a note of the file extension. If it is `.png` we will use `png` packa

```{r}
 economics %>% 
  mutate(unemploy_per = unemploy / pop) %>% 
  ggplot(aes(date, unemploy_per)) +
  annotation_raster(readPNG("UTSA_Logo_Website_UTSA_Logo_Stacked_Center.png"),
                    xmin = -Inf, xmax = Inf, ymin = 0.015, ymax = 0.05, interpolate = T) +
  geom_line(color = "#69be0f", size = 1) +
  scale_y_continuous(labels = scales::percent)

```


```{r}
 economics %>% 
  mutate(unemploy_per = unemploy / pop) %>% 
  ggplot(aes(date, unemploy_per)) +
  annotation_raster(readPNG("UTSA_Logo_Website_UTSA_Logo_Stacked_Center.png"),
                    xmin = as.Date("2000-01-01"), xmax = as.Date("2015-01-01"), 
                    ymin = 0.015, ymax = 0.02, interpolate = T) +
  geom_line(color = "#69be0f", size = 1) +
  scale_y_continuous(labels = scales::percent)

```

## Use custom fonts with showtext

Finally, let's use custom fonts using `showtext` package.

```{r}
font_add_google("Akronim", "akronim")
showtext_auto()
```


```{r fig.showtext = TRUE}
p2 +
  labs(title = "Monthly Unemployment Rate") +
  theme(plot.title = element_text(family = "akronim", size = 16))
```





